---
title: "Onyeka Isamah, oi2142, Homework 2"
output: github_document
---


## Problem 1, Mr Trash Wheel Dataset:
  
  #Reading and cleaning Mr. Trash Wheel data set using read_excel and janitor::clean_names()
  
  #Omitting rows that do not include dumpster specific data
  
```{r setup}
library(tidyverse)
library (readxl)
library(skimr)
```


```{r Reading Cleaning Dropping}
trash_wheel = 
  read_excel("./data/NEW_mr_trash_wheel", sheet = "Mr. Trash Wheel", range = "a2:n408")
trash_wheel = 
    janitor::clean_names(trash_wheel)
trash_wheel = drop_na(trash_wheel)
```
  # Rounding the number of sports balls to nearest integer

```{r Rounding}
trash_wheel$sports_balls <- round(trash_wheel$sports_balls)
trash_wheel
```
 # Reading & Cleaning Precipitation 2018

```{r Cleaning precipitation data for 2018}
precipitation_2018 = 
  read_excel("./data/mr_trash_wheel.xlsx", sheet = "2018 Precipitation", range = "a2:B14")

precipitation_2018 = 
  janitor::clean_names(precipitation_2018)

precipitation_2018$year <- 2018
  
precipitation_2018  
```


 #Reading & Cleaning Precipitation 2019

```{r Cleaning precipitation data for 2019}
precipitation_2019 = 
  read_excel("./data/mr_trash_wheel.xlsx", sheet = "2019 Precipitation", range = "a2:b8") 

precipitation_2019 = 
  janitor::clean_names(precipitation_2019)

precipitation_2019$year <- 2019
  precipitation_2019  

```

 #Combining precipitation datasets 

```{r Combining / Merging data sets}
combined_precipitation <- bind_rows(precipitation_2018, precipitation_2019)
                                   
combined_precipitation<- mutate (combined_precipitation, MonthName = month.name[month])

combined_precipitation
```



The Mr. Trashwheel dataset has `r nrow(trash_wheel)` observations and `r ncol(trash_wheel)` variables. The variables account for each `dumpster` number, `weight`, and `volume` as well as the `month`, `year`, and `date`. The dataset also includes variables of the type of trash: `plastic`, `polystyrene`, `cigarette_butts`, `glass_bottles`, `grocery_bags`, `chip_bags`, and  `sports_balls`. 



The Precipiation data set has `r nrow(combined_precipitation)` observations for the amount of rainfall that occurred and `r ncol(combined_precipitation)` variables - which were `month`, `year`, and the `total` amount of rainfall.



Find sum and median**

### Problem 2 - FiveThirtyEight Dataset

 ###Clean the data in pols-month.csv. 

```{r Cleaning pols_df}
pols_df = 
  read_csv(file = "./data/fivethirtyeight_datasets/pols-month.csv") %>% 
  
  separate(mon, c("year", "month", "day")) %>% 
  
  mutate(
    month = month.name[as.numeric(month)],

    president = recode(prez_dem, "0" = "gop", "1"="dem"))   
  
    pols_df = select(pols_df, -prez_dem, -prez_gop, -day)
```
 ### Cleaning the snp_csv file

```{r cleaning snp_df}
snp_df = 
  read_csv(file = "./data/fivethirtyeight_datasets/snp.csv")

  janitor::clean_names(snp_df)  

snp_df=  separate(snp_df, "date", into =  c("month", "day", "year"), convert = TRUE) %>% 
  mutate( 
    year, year = if_else( year<= 15, year + 2000, year + 1900)) %>% 
      arrange(year, month, .by_group = TRUE) %>% 
 mutate(
    year = as.character(year),  
      month = month.name[as.numeric(month)], 
      month = str_to_lower(month))

```

 ### Tidying unemployment data, before merging
```{r Tidying unemployment_df}
unemployment_df = 
  read_csv("./data/fivethirtyeight_datasets/unemployment.csv") %>% 

  janitor::clean_names() %>% 
  
  pivot_longer(jan:dec,
    names_to = "month", 
    values_to = "unemployment") %>% 

  mutate(
  
  month = recode(month, 
    jan = "january", feb = "february", mar = "march", apr = "april", may = "may", jun = "june", jul = "july", aug =  "august", sep = "september", nov = "november", dec = "december"),
  
    year = as.character(year)
)
```

### Merging 
##Merging pols_df and snp_df (left join), then merging the result to the unemployment_df
```{r}
snp_pol_combo = left_join(pols_df,snp_df)

five_thirty_eight_joined = left_join(snp_pol_combo, unemployment_df) 
```


### Describing the data


