---
title: "Onyeka Isamah, oi2142, Homework 2"
output: github_document
---


## Problem 1, Mr Trash Wheel Dataset:
  
  #Reading and cleaning Mr. Trash Wheel data set using read_excel and janitor::clean_names()
  
  #Omitting rows that do not include dumpster specific data
  
```{r setup}
library(tidyverse)
library (readxl)
library(skimr)
```


```{r Reading Cleaning Dropping}
trash_wheel = 
  read_excel("./data/NEW_mr_trash_wheel", sheet = "Mr. Trash Wheel", range = "a2:n408")
trash_wheel = 
    janitor::clean_names(trash_wheel)
trash_wheel = drop_na(trash_wheel)
```
  # Rounding the number of sports balls to nearest integer

```{r Rounding}
trash_wheel$sports_balls <- round(trash_wheel$sports_balls)
trash_wheel
```
 # Reading & Cleaning Precipitation 2018

```{r}
precipitation_2018 = 
  read_excel("./data/mr_trash_wheel.xlsx", sheet = "2018 Precipitation", range = "a2:B14")

precipitation_2018 = 
  janitor::clean_names(precipitation_2018)

precipitation_2018$year <- 2018
  
precipitation_2018  
```


 #Reading & Cleaning Precipitation 2019

```{r}
precipitation_2019 = 
  read_excel("./data/mr_trash_wheel.xlsx", sheet = "2019 Precipitation", range = "a2:b8") 

precipitation_2019 = 
  janitor::clean_names(precipitation_2019)

precipitation_2019$year <- 2019
  precipitation_2019  
```

 #Combining precipitation datasets 

```{r}
combined_precipitation <- bind_rows(precipitation_2018, precipitation_2019)
                                   
combined_precipitation<- mutate (combined_precipitation, MonthName = month.name[month])
```


The data set for Precipitation in 2019 has `r nrow(precipitation_2019)` observations for the total amount of rainfall that occurred in the year 2019 and `r ncol(precipitation_2019)` variables - which were the month, the total amount of rainfall and the year. 

The dataset for Precipitation in 2018 has `r nrow(precipitation_2018)` observations for the total amount of rainfall that occurred in the year 2018 and `r ncol(precipitation_2018)` variables - which were the month, the total amount of rainfall and the year. 

`r colSums(combined_precipitation, total)`
na.rm = TRUE

