---
title: "Onyeka Isamah, oi2142, Homework 2"
output: github_document
---


### Problem 1, Mr Trash Wheel Dataset:
  
##### Reading and cleaning Mr. Trash Wheel data set using read_excel and janitor::clean_names()
  
##### Omitting rows that do not include dumpster specific data
  
```{r setup}
library(tidyverse)
library (readxl)
library(skimr)
```


```{r Reading Cleaning Dropping}
trash_wheel = 
  read_excel("./data/NEW_mr_trash_wheel", sheet = "Mr. Trash Wheel", range = "a2:n408")
trash_wheel = 
    janitor::clean_names(trash_wheel)
trash_wheel = drop_na(trash_wheel)
```
##### Rounding the number of sports balls to nearest integer

```{r Rounding}
trash_wheel$sports_balls <- round(trash_wheel$sports_balls)
trash_wheel
```
##### Reading & Cleaning Precipitation 2018

```{r Cleaning precipitation data for 2018}
precipitation_2018 = 
  read_excel("./data/mr_trash_wheel.xlsx", sheet = "2018 Precipitation", range = "a2:B14")

precipitation_2018 = 
  janitor::clean_names(precipitation_2018)

precipitation_2018$year <- 2018
  
precipitation_2018  
```


##### Reading & Cleaning Precipitation 2019

```{r Cleaning precipitation data for 2019}
precipitation_2019 = 
  read_excel("./data/mr_trash_wheel.xlsx", sheet = "2019 Precipitation", range = "a2:b8") 

precipitation_2019 = 
  janitor::clean_names(precipitation_2019)

precipitation_2019$year <- 2019
  precipitation_2019  
  

```

##### Combining precipitation datasets 

```{r Combining / Merging data sets}
combined_precipitation <- bind_rows(precipitation_2018, precipitation_2019)
                                   
combined_precipitation<- mutate (combined_precipitation, MonthName = month.name[month])

combined_precipitation


```



The Mr. Trashwheel dataset has `r nrow(trash_wheel)` observations and `r ncol(trash_wheel)` variables. The variables account for each `dumpster` number, `weight`, and `volume` as well as the `month`, `year`, and `date`. The dataset also includes variables of the type of trash: `plastic`, `polystyrene`, `cigarette_butts`, `glass_bottles`, `grocery_bags`, `chip_bags`, and  `sports_balls`. The median number of `sports_balls` in `year` 2019 is `r median(trash_wheel$sports_balls, year = 2017)`



The Precipitation data set has `r nrow(combined_precipitation)` observations for the amount of rainfall that occurred and `r ncol(combined_precipitation)` variables - which were `month`, `year`, and the `total` amount of rainfall. The total amount of precipitation in 2018 is `r sum(precipitation_2018$total)`.



### Problem 2 - FiveThirtyEight Dataset

##### Clean the data in pols-month.csv. 

```{r Cleaning pols_df}
pols_df = 
  read_csv(file = "./data/fivethirtyeight_datasets/pols-month.csv") %>% 
  
  separate(mon, c("year", "month", "day")) %>% 
  
  mutate(
    month = month.name[as.numeric(month)],

    president = recode(prez_dem, "0" = "gop", "1"="dem"))   
  
    pols_df = select(pols_df, -prez_dem, -prez_gop, -day)
    pols_df
```
##### Cleaning the snp_csv file

```{r cleaning snp_df}
snp_df = 
  read_csv(file = "./data/fivethirtyeight_datasets/snp.csv")

  janitor::clean_names(snp_df)  

snp_df=  separate(snp_df, "date", into =  c("month", "day", "year"), convert = TRUE) %>% 
  mutate( 
    year, year = if_else( year<= 15, year + 2000, year + 1900)) %>% 
      arrange(year, month, .by_group = TRUE) %>% 
 mutate(
    year = as.character(year),  
      month = month.name[as.numeric(month)], 
      month = str_to_lower(month))
snp_df
```

##### Tidying unemployment data, before merging
```{r Tidying unemployment_df}
unemployment_df = 
  read_csv("./data/fivethirtyeight_datasets/unemployment.csv") %>% 

  janitor::clean_names() %>% 
  
  pivot_longer(jan:dec,
    names_to = "month", 
    values_to = "unemployment") %>% 

  mutate(
  
  month = recode(month, 
    jan = "january", feb = "february", mar = "march", apr = "april", may = "may", jun = "june", jul = "july", aug =  "august", sep = "september", nov = "november", dec = "december"),
  
    year = as.character(year)
)
unemployment_df
```

##### Merging 
###### Merging pols_df and snp_df (left join), then merging the result to the unemployment_df
```{r}
snp_pol_combo = left_join(pols_df,snp_df)

five_thirty_eight_joined = left_join(snp_pol_combo, unemployment_df) 
five_thirty_eight_joined
```


##### Describing the data


The Pols_df data set has `r nrow(pols_df)` observations  and `r ncol(pols_df)` variables related to politicians who were democratic or republican between the years 1947 and 2015.

The snp_df data set has  `r nrow(snp_df)` observations  and `r ncol(snp_df)` variables related to Standard & Poor’s stock market index (S&P) from 1950 - 2015. The variables were `month`, `year`, and `close`. 

The unemployment_df data set has  `r nrow(unemployment_df)` observations  and `r ncol(unemployment_df)` variables related to the percentage of unemployment by `month` (January- February) from the years 1948-2015.

The resulting data set that joined pols,snp, and unemployment (five_thirty_eight_joined) has `r nrow(five_thirty_eight_joined)` observations  and `r ncol(five_thirty_eight_joined)` variables. 

 
 
##### Problem3 NYC Open Data

 Cleaning the data set
```{r cleaning babynames data set}
babyname_df = distinct(
  read_csv("./data/Popular_baby_names.csv") %>% 
  janitor::clean_names() %>% 
  
  mutate(ethnicity = recode(ethnicity,
   "WHITE NON HISP" = "WHITE NON HISPANIC", 
   "WHITE NON HISPANIC"="WHITE NON HISPANIC",  
   "BLACK NON HISP" =  "BLACK NON HISPANIC",  
   "BLACK NON HISPANIC" = "BLACK NON HISPANIC",
 "ASIAN AND PACI" ="ASIAN AND PACIFIC ISLANDER",  
  "ASIAN AND PACIFIC ISLANDER"= "ASIAN AND PACIFIC ISLANDER"
)) %>% 
  
mutate(
  year_of_birth = as.character(year_of_birth)))
```

##### Popularity of the name *Olivia* as a female baby name over the years

```{r Olivia Ranking}
 olivia_rank=(select(babyname_df, childs_first_name, year_of_birth, gender, rank, ethnicity, -count) %>% 
                     
    filter(childs_first_name == "OLIVIA", 
    gender == "FEMALE") %>% 
  
      pivot_wider(
        names_from = "year_of_birth",
        values_from = "rank") %>% 
        group_by(ethnicity))
olivia_rank
```

##### Most popular baby names for male children over time

```{r Male Ranking}
 male_rank= pivot_wider(
   
  (select(babyname_df, -count) %>% 
  filter(gender == "MALE") %>% 
  pivot_wider(
  names_from = "year_of_birth", 
  values_from = "rank") %>% 
  group_by(ethnicity)))
```


###### Scatter Plot
Creating data frame to be plotted, for white non hispanic males who were born in 2016
```{r}
whitemale_plot = filter(babyname_df, gender == "MALE", ethnicity == "WHITE NON HISPANIC", year_of_birth == 2016)
```

Scatterplot for white non hispanic males who were born in 2016

```{r scatter plot}
ggplot(whitemale_plot, aes(x = rank, y = count)) + 
  geom_point()
```
